---
title: "Intro to Differential Privacy"
author: "Caleb Moses"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  rmarkdown::html_document:
    theme: lumen
logo: images/dragonfly-logo-rgb.png
css: dragonfly.css
font-family: 'Helvetica'
---

<script type="text/javascript" src="http://livejs.com/live.js"></script>

<style>
pre code, pre, code {
  white-space: pre !important;
  overflow-x: scroll !important;
  word-break: keep-all !important;
  word-wrap: initial !important;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
```

```{r}
library(tidyverse)
library(kableExtra)
library(reticulate)
```

## Overview

- What is Differential Privacy?
- Examples of Differentially Private Mechanisms

## Simple scenario

> Alice permits her medical data to be used by third parties for medical research.
>
> Her insurance company conducts a study with the medical data, which concludes smoking is bad for health
>
> Her insurance company asks customers whether they smoke during the application process, and use this
information to justify raising Alice's premiums

Has Alice been harmed by allowing her data to be used by third parties

## What is Differential Privacy?

Differential Privacy is a mathematical abstraction of disclosure risk which argues that the goal of 
confidentiality should be to control the likelihood of distinguishing


























